---
title: "Simulation Blocks 1A-1E for NCA With Psychological Data"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
execute:
  echo: true
  warning: false
  message: false
---

## Overview

This notebook runs the first five one-factor-at-a-time simulation blocks for the NCA methods paper.

The logic is the following:

- **1A**: vary score construction only;
- **1B**: vary reliability/loading strength only;
- **1C**: vary ordinal scaling only, including the optional threshold-spacing manipulation;
- **1D**: vary latent skewness only;
- **1E**: vary range restriction only.

The baseline latent model is a simple linear relation without a true latent necessity mechanism.

## Load functions

```{r}
source("nca_simulation_1A_1E.R")
```

## Global settings

For local testing, keep the number of replications low. For the paper, increase `reps_main` to something like 500.

```{r}
reps_demo <- 25
n_main <- 1000
test_rep_main <- 200
```

# Simulation 1A: Score construction

## Rationale

This block keeps the latent relation, measurement model, and sample fixed. It compares NCA results obtained from:

- sum scores,
- mean scores,
- CFA factor scores.

## Run

```{r}
res_1A <- simulate_1A_score_construction(
  reps = reps_demo,
  n = n_main,
  test_rep = test_rep_main,
  progress = FALSE
)

sum_1A <- summarise_simulation_results(res_1A)
sum_1A
```

## Plot

```{r}
plot_effect_size_density(res_1A)
```

# Simulation 1B: Reliability / loading strength

## Rationale

This block varies item loading strength while keeping everything else fixed.

## Run

```{r}
res_1B <- simulate_1B_reliability(
  reps = reps_demo,
  loading_levels = c(0.50, 0.70, 0.85),
  n = n_main,
  test_rep = test_rep_main,
  progress = FALSE
)

sum_1B <- summarise_simulation_results(res_1B)
sum_1B
```

## Plot

```{r}
plot_mean_effects(sum_1B, "loading")
```

# Simulation 1C: Ordinal scaling

## Rationale

This block varies:

- number of response categories (5 vs 7), and
- threshold spacing (**symmetric** vs **skewed**).

The second factor implements the optional extension we discussed.

## Run

```{r}
res_1C <- simulate_1C_ordinal_scaling(
  reps = reps_demo,
  n_cat_levels = c(5, 7),
  threshold_spacings = c("symmetric", "skewed"),
  n = n_main,
  test_rep = test_rep_main,
  progress = FALSE
)

sum_1C <- summarise_simulation_results(res_1C)
sum_1C
```

## Plot

```{r}
plot_effect_size_density(res_1C, facet_var = "threshold_spacing")
```

# Simulation 1D: Skewness

## Rationale

This block keeps the linear latent relation fixed while changing the shape of the latent distributions.

## Run

```{r}
res_1D <- simulate_1D_skewness(
  reps = reps_demo,
  n = n_main,
  test_rep = test_rep_main,
  progress = FALSE
)

sum_1D <- summarise_simulation_results(res_1D)
sum_1D
```

## Plot

```{r}
plot_effect_size_density(res_1D, facet_var = "skew_label")
```

# Simulation 1E: Range restriction

## Rationale

This block varies sample composition while leaving the latent relation and measurement structure unchanged.

## Run

```{r}
res_1E <- simulate_1E_range_restriction(
  reps = reps_demo,
  restrictions = c("full", "x_upper", "x_middle", "y_upper"),
  n = n_main,
  test_rep = test_rep_main,
  progress = FALSE
)

sum_1E <- summarise_simulation_results(res_1E)
sum_1E
```

## Plot

```{r}
plot_effect_size_density(res_1E, facet_var = "restriction")
```

# Save outputs

```{r}
# Uncomment when you want to save the results locally.
# save_simulation_bundle(res_1A, "res_1A")
# save_simulation_bundle(res_1B, "res_1B")
# save_simulation_bundle(res_1C, "res_1C")
# save_simulation_bundle(res_1D, "res_1D")
# save_simulation_bundle(res_1E, "res_1E")
```

# Next steps

For the main paper, the natural next step after these one-factor blocks is the reduced factorial design.
