---
title: "Supplementary Materials"
subtitle: "Necessary Condition Analysis With Psychological Data: A Psychometric Critique"
author: "Tommaso Feraco"
format:
  html:
    toc: true
    number-sections: true
    df-print: paged
execute:
  echo: false
  warning: false
  message: false
  cache: false
---

```{r}
library(dplyr)
library(tidyr)
library(knitr)

results <- read.csv(here::here("simulations/simulation_results.csv"))
results <- read.csv(here::here("simulations/results_table.csv"))
results_n <- read.csv(here::here("simulations/simulation_results_n.csv"))
results_n <- read.csv(here::here("simulations/results_table_n.csv"))

fmt_num <- function(x, digits = 3) round(x, digits)
```

# Overview

This supplement reports the most useful numerical summaries from the simulations.  
Across tables, the main quantities are:

- `Observed d`: mean observed NCA effect size across replications.
- `SD(d)`: standard deviation of the observed NCA effect size across replications.
- `Mean r(X,Y)`: mean observed correlation between X and Y across replications.
- `X skew` and `Y skew`: mean observed skewness of the analyzed X and Y scores across replications.
- `Delta from latent base`: mean difference between the observed effect and the latent-reference NCA effect from the full latent sample in the same design cell.
- `Delta from latent selected`: mean difference between the observed effect and the condition-specific latent reference after selection. This column is only used for the range-restriction block.

```{r}
block_labels <- c(
  "1A_score_construction" = "Score construction",
  "1B_reliability" = "Reliability / loading",
  "1C_response_categories" = "Response categories",
  "1D_threshold_shift" = "Threshold shift",
  "1E_range_restriction" = "Range restriction"
)

tbl_overview <- results %>%
  mutate(Block = recode(block, !!!block_labels)) %>%
  group_by(Block) %>%
  summarise(
    `Mean observed d` = fmt_num(mean(mean_effect_size, na.rm = TRUE)),
    `Mean SD(d)` = fmt_num(mean(sd_effect_size, na.rm = TRUE)),
    `Mean r(X,Y)` = fmt_num(mean(mean_correlation, na.rm = TRUE)),
    `Mean X skew` = fmt_num(mean(mean_x_skew_obs, na.rm = TRUE)),
    `Mean Y skew` = fmt_num(mean(mean_y_skew_obs, na.rm = TRUE)),
    `Mean delta from latent base` = fmt_num(mean(mean_delta_from_base_full, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  arrange(Block)

kable(tbl_overview, caption = "Table S1. Pooled overview of the five simulation blocks.")
```

# Latent-reference effects for the main design cells

```{r}
tbl_latent_reference <- results %>%
  group_by(`Sample size` = n_input, Beta = beta) %>%
  summarise(
    `Latent reference d` = fmt_num(mean(mean_latent_effect_base_full, na.rm = TRUE)),
    `Mean latent r(X,Y)` = fmt_num(mean(mean_correlation, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  arrange(`Sample size`, Beta)

kable(tbl_latent_reference, caption = "Table S2. Latent-reference NCA effect sizes for the main design cells.")
```

# Score construction

```{r}
tbl_score_construction <- results %>%
  filter(block == "1A_score_construction") %>%
  transmute(
    `Sample size` = n_input,
    Beta = beta,
    `Score type` = score_type,
    `Observed d` = fmt_num(mean_effect_size),
    `SD(d)` = fmt_num(sd_effect_size),
    `Mean r(X,Y)` = fmt_num(mean_correlation),
    `X skew` = fmt_num(mean_x_skew_obs),
    `Y skew` = fmt_num(mean_y_skew_obs),
    `Delta from latent base` = fmt_num(mean_delta_from_base_full)
  ) %>%
  arrange(`Sample size`, Beta, `Score type`)

kable(tbl_score_construction, caption = "Table S3. Score-construction block.")
```

# Reliability / loading strength

```{r}
tbl_reliability <- results %>%
  filter(block == "1B_reliability") %>%
  mutate(Loading = sub("loading_", "", condition_label)) %>%
  transmute(
    `Sample size` = n_input,
    Beta = beta,
    Loading,
    `Observed d` = fmt_num(mean_effect_size),
    `SD(d)` = fmt_num(sd_effect_size),
    `Mean r(X,Y)` = fmt_num(mean_correlation),
    `X skew` = fmt_num(mean_x_skew_obs),
    `Y skew` = fmt_num(mean_y_skew_obs),
    `Delta from latent base` = fmt_num(mean_delta_from_base_full)
  ) %>%
  arrange(`Sample size`, Beta, Loading)

kable(tbl_reliability, caption = "Table S4. Reliability / loading block.")
```

# Response categories

```{r}
tbl_categories <- results %>%
  filter(block == "1C_response_categories") %>%
  mutate(`Response categories` = sub("n_cat_", "", condition_label)) %>%
  transmute(
    `Sample size` = n_input,
    Beta = beta,
    `Response categories`,
    `Observed d` = fmt_num(mean_effect_size),
    `SD(d)` = fmt_num(sd_effect_size),
    `Mean r(X,Y)` = fmt_num(mean_correlation),
    `X skew` = fmt_num(mean_x_skew_obs),
    `Y skew` = fmt_num(mean_y_skew_obs),
    `Delta from latent base` = fmt_num(mean_delta_from_base_full)
  ) %>%
  arrange(`Sample size`, Beta, `Response categories`)

kable(tbl_categories, caption = "Table S5. Response-category block.")
```

# Threshold shift

```{r}
tbl_threshold_shift <- results %>%
  filter(block == "1D_threshold_shift") %>%
  separate(condition_label, into = c("Direction", "Magnitude"), sep = "_") %>%
  transmute(
    `Sample size` = n_input,
    Beta = beta,
    Direction,
    Magnitude,
    `Mean r(X,Y)` = fmt_num(mean_correlation),
    `X skew` = fmt_num(mean_x_skew_obs),
    `Y skew` = fmt_num(mean_y_skew_obs),
    `Observed d` = fmt_num(mean_effect_size),
    `SD(d)` = fmt_num(sd_effect_size),
    `Delta from latent base` = fmt_num(mean_delta_from_base_full)
  ) %>%
  arrange(`Sample size`, Beta, Direction, as.numeric(Magnitude))

kable(tbl_threshold_shift, caption = "Table S6. Threshold-shift block, including the resulting observed skewness of X and Y.")
```

# Range restriction

```{r}
tbl_range_restriction <- results %>%
  filter(block == "1E_range_restriction") %>%
  transmute(
    `Sample size` = n_input,
    Beta = beta,
    Restriction = condition_label,
    `Latent d (selected sample)` = fmt_num(mean_latent_effect_condition_selected),
    `Selection effect on latent d` = fmt_num(mean_selection_effect_on_latent_d),
    `Observed d` = fmt_num(mean_effect_size),
    `SD(d)` = fmt_num(sd_effect_size),
    `Mean r(X,Y)` = fmt_num(mean_correlation),
    `X skew` = fmt_num(mean_x_skew_obs),
    `Y skew` = fmt_num(mean_y_skew_obs),
    `Delta from latent selected` = fmt_num(mean_delta_from_condition_selected)
  ) %>%
  arrange(`Sample size`, Beta, Restriction)

kable(tbl_range_restriction, caption = "Table S7. Range-restriction block. The key comparison is between the observed effect and the latent reference in the selected sample.")
```

# Latent-only sample-size simulation

```{r}
tbl_latent_n <- results_n %>%
  transmute(
    `Sample size` = n_input,
    Beta = beta,
    `Observed d` = fmt_num(mean_effect_size),
    `SD(d)` = fmt_num(sd_effect_size),
    `95% CI low` = fmt_num(ci_low),
    `95% CI high` = fmt_num(ci_high),
    `Mean r(X,Y)` = fmt_num(mean_correlation),
    `X skew` = fmt_num(mean_x_skew),
    `Y skew` = fmt_num(mean_y_skew)
  ) %>%
  arrange(`Sample size`)

kable(tbl_latent_n, caption = "Table S8. Latent-only sample-size simulation used to evaluate how the NCA effect changes as N increases while the latent relation remains fixed.")
```
